---
title: Analysis and figures for 'Estimating bioturbation from replicated small-sample
  radiocarbon ages'
author: "Andrew M. Dolman"
date: "05/03/2021"
output:
  bookdown::pdf_document2:
    toc: no
    keep_md: yes
    fig_crop: no
  html_document:
    fig_dev: cairo_png
subtitle: Revised
params:
  doc_section: analysis
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message = FALSE, warning = FALSE, cache = TRUE,
                      autodep = TRUE, tidy = FALSE, dpi = 300)

library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(Bchron)
library(ggridges)
library(egg)
library(sp)
library(kableExtra)
library(readr)
library(knitr)
library(rnaturalearth)
library(splines)
library(directlabels)
```


## Load all radiocarbon data


```{r key}
key <- tibble(core = c("GeoB 10054-4", "GeoB 10066-7", "GeoB 10058-1",
                       "SO213-84-2", "SO213-84-2", "SO213-84-2",
                       "OR1-1218-C2-BC"),
              cruise = c("SO184","SO184","SO184", "SO213/2", "SO213/2",
                         "SO213/2", "OR1-1218"),
              site = c("GeoB 10054-4/58-1", "GeoB 10066-7", "GeoB 10054-4/58-1",
                       "SO213-84-2", "SO213-84-2", "SO213-84-2",
                       "OR1-1218-C2-BC"))
```



```{r core.info}
core.info <- structure(list(
  core = c(
    "GeoB 10054-4", "GeoB 10066-7", "OR1-1218-C2-BC",
    "GeoB 10058-1", "SO213-84-2", "SO213-84-2", "SO213-84-2"
  ), site = c(
    "GeoB 10054-4/58-1",
    "GeoB 10066-7", "OR1-1218-C2-BC", "GeoB 10054-4/58-1", "SO213-84-2",
    "SO213-84-2", "SO213-84-2"
  ), Latitude = c(
    -8.68166666666667,
    -9.39266666666667, 10.9005, -8.683, -45.1166666666667, -45.1166666666667,
    -45.1166666666667
  ), Longitude = c(
    112.668333333333, 118.5755,
    115.307666666667, 112.639, 174.583333333333, 174.583333333333,
    174.583333333333
  ), Waterdepth_m = c(
    1076, 1635, 2208, 1103, 992,
    992, 992
  ), Longitude.char = c(
    "112°40.10'E", "118°34.53'E", "115°18.46'E",
    "112°38.34'E", "174°35'11.4\\E", "174°35'11.4\\E", "174°35'11.4\\E"
  ), Latitude.char = c(
    "8°40.90'S", "9°23.56'S", "10°54.03'N",
    "8°40.98'S", "45°7'28.2\\S", "45°7'28.2\\S", "45°7'28.2\\S"
  ),
  approx.sed.acc.rate.cm.ka = c(18, 13, 2, 18, 3.8, 3.8, 3.8), cruise = c(
    "SO184", "SO184", "OR1-1218", "SO184", "SO213/2",
    "SO213/2", "SO213/2"
  )
), row.names = c(NA, -7L), class = c(
  "tbl_df",
  "tbl", "data.frame"
))
```



## Load data

```{r read-csv-files}
# read in data
dat.small.n <- read.csv("small-n-data.csv", stringsAsFactors = FALSE, header = TRUE) %>%
  mutate(`Size fraction [$\\mu$m]` = size.fraction) %>%
  mutate(n.forams.grp = factor(n.forams.grp,
    ordered = TRUE,
    levels = c("1", "3", "5", "5-6", "10", "30")
  )) 


dat.downcore <- read.csv("downcore-data.csv", stringsAsFactors = FALSE, header = TRUE) %>% 
  mutate(`Size fraction [$\\mu$m]` = size.fraction) %>% 
  as_tibble()
```

## Summary tables

```{r  tab-site-summary}
site.summary <- core.info %>% 
  ungroup() %>% 
  mutate(Latitude.char = as.character(sp::dd2dms(Latitude, NS = TRUE)),
         Longitude.char = as.character(sp::dd2dms(Longitude, NS = FALSE))) %>% 
  mutate(Latitude.char = gsub("d", "$^{\\circ}$", Latitude.char),
         Longitude.char = gsub("d", "$^{\\circ}$", Longitude.char)) %>%
  dplyr::select(Core = core, Cruise = cruise,
        Latitude = Latitude.char, Longitude = Longitude.char,
         `Water depth [m]` = Waterdepth_m) %>% 
  distinct()

site.summary %>% 
  knitr::kable(., "latex", booktabs = TRUE, caption = "Sampled locations", escape = FALSE, linesep = "")
```

```{r summarise-dataset}
dat.small.n.smry <- dat.small.n %>%
  group_by(core, site, cruise, subcore, species, size.fraction,
           `Size fraction [$\\mu$m]`, depth.range, n.forams.grp) %>%
  dplyr::summarise(n.reps = n()) %>%
  filter(n.reps > 1) %>% 
  ungroup()

dat.combined <- bind_rows(dat.small.n.smry, dat.downcore) %>% 
  dplyr::select(core, subcore, cruise, 
         depth.range, species,
         `Size fraction [$\\mu$m]`, n.forams.grp, n.reps) %>% 
  filter(grepl("x", n.forams.grp, fixed = TRUE) == FALSE)


dat.summary <- dat.combined %>%
  dplyr::select(core, subcore, cruise,
         depth.range, species,
         `Size fraction [$\\mu$m]`, n.forams.grp, n.reps) %>%
  mutate(n.forams.grp = gsub(">", "$>$", n.forams.grp)) %>% 
  dplyr::rename(Core = core,
         Cruise = cruise,
         `Depth [cm]` = depth.range,
         Species = species,
         `$\\text{n}_{f}$` = n.forams.grp,
         `$\\text{n}_{rep}$` = n.reps) %>%
  rowwise() %>%
  mutate(subcore = readr::parse_number(subcore),
         Core = ifelse(is.na(subcore) == FALSE, paste0(Core, " (", subcore, ")", collapse = ","), Core)) %>% 
  dplyr::select(-subcore) %>% 
  arrange(Core, `Depth [cm]`) %>%
  mutate(Species = kableExtra::cell_spec(Species, italic = TRUE)) 
```


```{r tab-data-summary}
dat.summary %>% 
  dplyr::select(#-Latitude, -Longitude,
         -Cruise) %>% 
  #distinct() %>% 
  knitr::kable(., "latex", booktabs = TRUE, caption = "Data summary", escape = FALSE, linesep = "")
```

## Calculate sedimentation rates


### Additional figure not in manuscript
```{r fig_sed-rates}
p.sed.acc.rates <- dat.downcore %>% 
  ggplot(aes(x = depth_cm, y = age.14C.cal, group = core, shape = core)) +
  geom_pointrange(aes(ymax = age.14C.cal + age.14C.cal.se, ymin = age.14C.cal - age.14C.cal.se)) +
  geom_smooth(method = "lm") +
  facet_wrap(~site, scales = "free", ncol = 2) +
  theme_bw()

p.sed.acc.rates
```


```{r fit_lm_for_sed_rates}
sed.rates <- dat.downcore %>% 
  group_by(core, site) %>% 
  do({
    lm1 <- lm(age.14C.cal ~ depth_cm, data = .)
    sumr <- summary(lm1)
    
    data.frame(term = "depth_cm", 
               estimate = sumr$coefficients[2,1], 
               std.error = sumr$coefficients[2,2])
  }) %>% 
  mutate(s = 1e03 * 1/estimate,
         s.err = (std.error/estimate) * s) %>% 
  dplyr::select(core, site, s, s.err) %>% 
  group_by(site) %>% 
  mutate(wt = (1/s.err) / sum(1/s.err)) %>% 
  mutate(s = sum(s * wt),
            s.err = sum(s.err * wt)) %>% 
  dplyr::select(-wt) %>% 
  ungroup() %>% 
  arrange(site, core)
```

```{r sed.rates_table}
sed.rates %>% 
  dplyr::select(site, s, s.err) %>% 
  distinct() %>% 
  arrange(site) %>% 
  knitr::kable(., digits = 2)
```

## Map of core sites

```{r map-palette}
map.info <- core.info %>% 
  dplyr::select(core, site, Longitude, Latitude) %>% 
  filter(core != "T86-10P") %>% 
  distinct() %>% 
  mutate(site = ifelse(site == "GeoB 10054-4/58-1", "GeoB 10054-4\nGeoB 10058-1", site))

# Hard code to remove dependency
#col.pal <- RColorBrewer::brewer.pal(6, name = "Set1")[c(1, 4, 5, 2, 3)]
col.pal <- c("#E41A1C", "#984EA3", "#FF7F00", "#377EB8", "#4DAF4A")
names(col.pal) <- map.info$site
```


```{r fig-map-pacific, fig.height=3.5, fig.width=6, fig.cap="Map of core sites"}
world <- rnaturalearth::ne_countries(scale = 'small', returnclass = 'sp')

worldmap2 <- ggplot(world, aes(x = long, y = lat, group = group)) +
  geom_polygon(colour = "NA", fill = "Darkgrey", show.legend = F) +
  coord_map("gall", 100,
            xlim = c(80, 185),
            ylim = c(-55, 25)) +
  geom_point(data = map.info,
             aes(Longitude, Latitude, colour = site, shape = site),
             inherit.aes = FALSE, show.legend = T,
             size = 3) +
  scale_color_manual("", values = col.pal) +
  scale_shape_manual("", values = c(15:18))+
  theme_bw() +
  theme(legend.position = "right") +
  guides(colour = guide_legend(ncol = 1)) +
  scale_x_continuous("Longitude", breaks = c(-90, 0, 90, 120, 150, 180, -180),
                     labels = function(x) paste0(x,"°")) +
  scale_y_continuous("Latitude", breaks = c(20, 0, -20, -40),
                     labels = function(x) paste0(x,"°")) 
worldmap2
```



## Calibration to calendar ages


```{r calibration_densities, warning=FALSE}
# Need the full densities so need to do calibrations here.
# Order to make later figure pretty
C14.for.cali <- dat.small.n %>%
  arrange(sample.label, n.forams.grp, age.14C) %>% 
  mutate(ID.no = 1:nrow(.))

calib <- prxytools::CalibrateAge(C14.for.cali,
                                 return.type = "lst",
                                 curve = "marine20")

# The summarised calendar ages are appended to the input data
C14 <- calib$df

# These are the full PDFs
cal.ages <- calib$cal.ages
```

```{r data_for_density_plots}
# create data.frames for plotting
C14.id <- C14 %>% 
  dplyr::select(ID.no, core, depth_cm, sample.label, n.forams.grp, age.14C.cal, age.14C.cal.se)

# convert odd Bchron format of full PDFs to something tidy
cali.pdf.df <- plyr::ldply(1:length(cal.ages), function(i){
  x <- cal.ages[[i]]
  if (is.na(x)==FALSE){data.frame(age = x[[1]]$ageGrid, d = x[[1]]$densities, .id = i)}else{
    data.frame(age = 0, d = 0, .id = i)
    }
  }) %>% 
  as_tibble() %>% 
  mutate(ID.no = as.numeric(.id)) %>% 
  # join to sample info
  left_join(C14.id, .) %>% 
  mutate(.id = as.numeric(.id))


cali.pdf.df.mean <- cali.pdf.df %>%
  dplyr::group_by(.id, ID.no, core, sample.label, n.forams.grp) %>%
  dplyr::summarise(mean.age.14C.cal = mean(age.14C.cal),
            max.d = max(d))%>% 
  ungroup() %>% 
  mutate(.id = as.numeric(.id))
```

```{r fig-cal-age-pdfs, fig.height=10, fig.width=8, fig.cap="Calendar age PDFs"}
# dont plot Lougheed data
  
p.cal.age <- cali.pdf.df %>% 
  filter(core != "T86-10P") %>% 
  mutate(.id = as.numeric(ID.no)) %>% 
  filter(is.na(d) == FALSE) %>% 
  ggplot(aes(x = age/1000, y = .id*max(d)/2, height = d,
             group = .id,
             fill = n.forams.grp)
  ) +
  geom_ridgeline() +
  geom_segment(data = filter(cali.pdf.df.mean, is.na(max.d) == FALSE, core != "T86-10P"), 
               aes(x = mean.age.14C.cal/1000,
                   xend = mean.age.14C.cal/1000,
                   y = (.id*max(max.d)/2),
                   yend = (.id*max(max.d)/2) + max.d,
                   linetype = "mean"),
               inherit.aes = F) +
  facet_wrap(~sample.label, scales = "free") +
  scale_y_continuous("") +
  scale_x_continuous("Calendar age [ka]") +
  theme_bw()+
  theme(axis.ticks.y = element_blank(), axis.text.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank()) 

p.cal.age
```


## Calibrated ages in MUC3_21-22


```{r MUC-age-pdfs, fig.height=4, fig.width=3}
f_labeller <- as_labeller(
  c("3" = "No. forams = 3", "5-6" = "No. forams = 5-6", "30" = "No. forams = 30")
)

p.cal.age.s <- cali.pdf.df %>% 
  filter(sample.label == "MUC3_21-22") %>% 
  mutate(.id = as.numeric(ID.no)) %>% 
  filter(is.na(d) == FALSE) %>% 
  ggplot(             ) +
  geom_ribbon(aes(x = age/1000, ymin = 0, ymax = d,
             group = .id), alpha = 0.5) +
  facet_wrap(~sample.label, scales = "free") +
  scale_y_continuous("") +
  scale_x_continuous("Calendar age [ka]") +
  theme_bw()+
  theme(axis.ticks.y = element_blank(), axis.text.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank()) +
  facet_grid(n.forams.grp~., labeller = f_labeller, as.table = FALSE)
```


```{r fig-MUC-age-pdfs, fig.height=4, fig.width=3, fig.cap=""}
p.cal.age.s
```


## Age ranges for text

```{r}
C14 %>% 
  filter(core == "SO213-84-2") %>% 
  group_by(subcore, n.forams.grp, depth_cm) %>% 
  summarise(sd.age = sd(age.14C.cal),
            age.min = min(age.14C.cal),
            age.max = max(age.14C.cal),
            range = paste(round(age.min/1000, 2), round(age.max/1000, 2)))
```



## Bias correction factors

```{r fig-F14C-bias-extreme, fig.width=4, fig.height=4, fig.cap=""}
# Pre-calculated bias correction factors are in file "sd.est.bias.RData". They
# can be recalculated with the code in "generate-bias-correction-factors.Rmd"
load("sd.est.bias.RData")

line.labs.a <- crossing(x = 2000, n = c(1, 2, 4))
 
line.labs.b <- crossing(x = 0.1, n = c(8, 16, 32, 1024)) 

line.labs <- bind_rows(line.labs.a, line.labs.b) %>% 
   mutate(y = sqrt(1/n),
         lab = paste0("italic(n[f])*\" = ", n, "\""))


fancy_sqrt <- function(l) {
  
     # turn in to character string in scientific notation
  nf <- 1/l^2
  
     l <- paste("sqrt(1/", nf,")")
     
     # return this as an expression
     parse(text=l)
}

clrs <- c("#1B9E77", "#D95F02", "#7570B3")

sd.est.bias %>% 
  filter(n.ind < 1025, L/s <= 11000,
         n.ind %in% c(64, 128, 256, 512) == FALSE) %>% 
  mutate(n.ind = factor(n.ind)) %>% 
  ggplot(aes(x = L/s, y = mean.sd.ind.ratio, group = n.ind)) +
  expand_limits(y = c(0, 1.05)) +
  coord_cartesian(xlim = c(0.1, 10000)) + 
  geom_point(alpha = 0.5, colour = clrs[3]) +
  geom_line(alpha = 1, colour = clrs[3]) +
  geom_line(aes(y = a), linetype = 3) +
  scale_x_continuous(trans = "log10",
                     breaks = c(0.1, 1, 10, 100, 1000, 10000),
                     labels = c(0.1, 1, 10, 100, 1000, 10000),
                     sec.axis = 
                       sec_axis(~10/., 
                                name = expression("Sediment accumulation rate [cm kyr"^-1*"]"),
                                breaks = c(0.01, 0.1, 1, 10, 100),
                     labels = c(0.01, 0.1, 1, 10, 100))) +
 
   scale_y_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1),
                      sec.axis = sec_axis(~., breaks = c(sqrt(1/c(1, 2, 4, 8, 16, 32, 1024))),
                                          labels = fancy_sqrt)) +
  geom_label(data = line.labs, aes(x = x, y = y, label = lab),
             group = NA,  colour = "black", hjust = 0, label.size = 0,
             parse = T) +
  geom_vline(xintercept = 10/1, linetype = 2, colour = clrs[2]) +
  annotation_logticks(sides = "bt") +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  labs(x = expression("True"~sigma[age]~"[kyr]"),
       y = expression("Estimated "~sigma[age]~"/"~"true"~sigma[age]))
```


```{r fig-F14C-bias-correction, fig.width=4, fig.height=4, fig.cap=""}
# Pre-calculated bias correction factors are in file "sd.est.bias.plausible.RData". They
# can be recalculated with the code in "generate-bias-correction-factors.Rmd"
load("sd.est.bias.plausible.RData")

lbls <- sd.est.bias.plausible %>% 
  group_by(n.ind) %>% 
  dplyr::summarise(y = max(1/mean.sd.ind.ratio)) %>% 
  mutate(x = Inf,
         lab = paste0("italic(n[f])*\" = ", n.ind, "\"")) %>% 
  filter(n.ind %in% c(1,2,3,4, 10, 1000))


fig.sd.pl.2 <- sd.est.bias.plausible %>% 
  mutate(n.ind = factor(n.ind)) %>% 
  filter(L/s <= 10/1) %>% 
  filter(n.ind %in% c(1,2,3,4, 10, 1000)) %>% 
  ggplot(aes(x = sd.age.ind.F14C, y = 1/mean.sd.ind.ratio, group = n.ind)) +
  geom_point(alpha = 0.5, colour = clrs[3])+
  geom_line(alpha = 1, colour = clrs[3]) +
  scale_x_continuous(trans = "log10")+
  annotation_logticks(sides = "b") +
  expand_limits(y = 1, x = 11000) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  scale_colour_discrete(expression(italic(n[f]))) +
  labs(x = expression("Estimated "~sigma[age]~" [years]"),
       y = "Bias correction factor") +
  geom_label(data = lbls, aes(x = x, y = y, label = lab),
             hjust = 1, label.size = 0, parse = T, size = 3)

fig.sd.pl.2
```



## Calculate age-variance and bioturbation depth


```{r calc-variances, message=FALSE, warning=FALSE}
SDofSD <- function(s, n) {

  if (any(n < 2, na.rm = TRUE))
    warning("n must be greater than or equal to 2")
  if (any(s < 0, na.rm = TRUE))
    warning("s must be >= 0")
  if(length(s) > 1 & length(n) == 1)
    n <- rep(n, length(s))
  if(length(n) > 1 & length(s) == 1)
    s <- rep(s, length(n))
  if(length(s) != length(n))
    stop("'s' and 'n' should be vectors of either the same length, ",
         "or one of them should be length == 1.")

  ifelse(n <= 300, {
    message("Method: Exact")
    g.a <- gamma((n - 1) / 2)
    g.b <- gamma(n / 2)
    s * (g.a / g.b) * sqrt((n - 1) / 2 - (g.b / g.a) ^ 2)
  }, {
    message("Method: Stirling's approximation")
    s * sqrt(exp(1) * (1 - 1 / n) ^ (n - 1) - 1)
  })
}


## 14C age and calendar age
C14.long <- C14 %>%
  # Remove 1 outlier
  filter(#is.outlier == FALSE,
         complete.cases(n.forams)) %>%
  
  #dplyr::select required columns
  dplyr::select(site, core, sample.label, depth_cm, n.forams.grp, n.forams, age.14C, age.14C.cal,
         age.14C.se, age.14C.cal.se, age.14C.count.se) %>%
  #pivot to long format
  gather(age.type, age, age.14C, age.14C.cal) %>%
  
  # associate correct age uncertainty, 14C or calibrated
  # could improve the pivoting to avoid this with the new tidyr::pivot_longer() 
  mutate(age.se = ifelse(age.type == "age.14C", age.14C.se, age.14C.cal.se)) %>% 
  dplyr::select(-age.14C.cal.se, -age.14C.se) %>%
  
  # group and calculate mean age at each depth
  group_by(site, sample.label, depth_cm, n.forams.grp, age.type) %>%
  mutate(mean.age = mean(age, na.rm = T)) %>% 
  ungroup()

C14.sd <- C14.long %>%
  group_by(site, core, age.type, sample.label, depth_cm, n.forams.grp) %>%
  dplyr::summarise(n.grps = n(),
            # mean should be taken on variance scale
            mean.machine.err = sqrt(mean(age.se^2, na.rm = T)),
            sd.age = sd(age),
            sd.age.corr = sqrt(sd.age^2 - ifelse(sample.label[1] != "GeoB 10066-7_53cm",
                                                mean.machine.err^2, (mean(age.14C.count.se))^2)),
            n.forams = mean(n.forams, na.rm = T),
            mean.age = mean(mean.age, na.rm = T),
            .groups = "keep")

# join new s estimates

sed.rates.sub <- sed.rates %>% 
  dplyr::select(-core) %>% 
  distinct()

C14.sd <- C14.sd %>% 
  left_join(., sed.rates.sub, by = "site")

# add a sedimentation rate for Lougheed
# use their published value
C14.sd <- C14.sd %>% 
  mutate(s = ifelse(core == "T86-10P", 2.2, s))

C14.sd <- C14.sd %>%
  mutate(sd.age.ind = sd.age.corr * n.forams^0.5,
         sd.of.sd = SDofSD(sd.age.corr, n.grps),
         sd.of.sd.ind = SDofSD(sd.age.ind, n.grps),
         sd.slice = sqrt(1/12 * (1000 * 1 / s)^2),
         est.biot.depth = sqrt(sd.age.ind^2 - sd.slice^2) * (s/1000),
         L = sqrt(sd.age.ind^2 - sd.slice^2) * (s/1000)) %>% 
  ungroup()
```

```{r get_correction_factors}
# Get correction factors
# These come from a fitted lm acting a lookup table
lm.plausible <- lm(log(sd.age.ind) ~ ns(log(sd.age.ind.F14C), 6) * as.factor(n.ind),
                   data = sd.est.bias.plausible)

coor.fac <- C14.sd %>%
  ungroup() %>% 
  filter(complete.cases(sd.age.ind),
         age.type == "age.14C") %>% 
  dplyr::select(site, sd.age.ind, n.forams, n.forams.grp) %>% 
  group_by(site, n.forams.grp) %>% 
  summarise_if(is.numeric, mean) %>% 
  mutate(n.forams = round(n.forams)) %>% 
  mutate(sd.age.ind.bc = exp(predict(lm.plausible,
                                     newdata = data.frame(sd.age.ind.F14C = sd.age.ind,
                                                          n.ind = n.forams)))) %>% 
  mutate(bias.correction = sd.age.ind.bc / sd.age.ind) %>% 
  ungroup() %>% 
  dplyr::select(site, n.forams.grp, bias.correction)


C14.sd <- left_join(C14.sd, coor.fac)

C14.sd <- C14.sd %>% 
  mutate(L.adj = L * bias.correction,
         sd.age.ind.bc = sd.age.ind * bias.correction)

C14.sd.ext <- crossing(C14.sd, n.forams.ext =
                         exp(seq(log(1), log(405), length.out = 100))) %>%
  mutate(sd.age.ext = sd.age.ind / n.forams.ext^0.5) %>%
  filter(core == "SO213-84-2")
```



## Check that SD declines as 1/sqrt(n)

Slope should be -1/2

```{r fig-SD-n-forams, fig.width=4.5, fig.height=3}
# Is the slope different to -0.5?
lm1 <- C14.sd %>% 
  filter(age.type == "age.14C.cal",
         core == "SO213-84-2") %>% 
lm(log10(sd.age.corr)~log10(n.forams), data = .)
summary(lm1)
confint(lm1)

sd.dec.dat <-  C14.sd %>% 
  filter(core == "SO213-84-2", age.type == "age.14C.cal") %>% 
  mutate(labl = gsub("_", " ", sample.label)) %>% 
  mutate(labl = gsub("cm", "", labl),
         labl = paste0(labl, " cm")) %>% 
  dplyr::select(labl, n.forams, sd.age.corr, sd.age)


p.sd.n.forams <- sd.dec.dat %>% 
  ggplot(aes(x = n.forams, y = sd.age.corr, group = NA)) +
  #ggplot(aes(x = n.forams, y = sd.age, group = NA)) +
  geom_abline(intercept = coef(lm1)[1], slope = -0.5,
              colour = "Red", linetype = 2) +
  geom_vline(xintercept = 1, colour = "Red", linetype = 2) +
  geom_smooth(method = "lm", colour = "grey28"#, fullrange = T
              ) +
  geom_point(aes(colour = labl, shape = labl)) +

  scale_x_continuous("No. forams per radiocarbon date",
                     trans = "log10", limits = c(1, 35)) +
  scale_y_continuous("SD age between samples", trans = "log10",
                     breaks = c(500, 1000, 2000, 3000, 4000, 5000)) +
  expand_limits(y = 5000) +
  annotation_logticks() +
  theme_bw() +
  scale_color_brewer("", type = "qual", palette = "Dark2") +
  scale_shape_discrete("") +
  theme(panel.grid.minor = element_blank())
p.sd.n.forams
```



## Age variance summary table

```{r var_summary}
age.var.summary <- C14.sd %>% 
  ungroup() %>% 
  filter(#age.type == "age.14C.cal",
         core != "T86-10P") %>% 
  dplyr::rename(total.age.sd = sd.age) %>% 
  dplyr::select(site, age.type, core, depth_cm, n.forams.grp, n.grps,
         total.age.sd, sd.of.sd, mean.machine.err, sd.slice, sd.age.corr,
         sd.age.ind, bias.correction, sd.age.ind.bc, sd.of.sd.ind, s, L, L.adj) 
```



```{r tab-age-var-summary}
options(knitr.kable.NA = "..")
age.var.summary %>% 
  filter(age.type == "age.14C.cal"#,
         ) %>% 
  dplyr::select(core, depth_cm, n.forams.grp, n.grps,
         total.age.sd, sd.of.sd,
         mean.machine.err, sd.age.ind,
         #sd.of.sd.ind,
         bias.correction,
         sd.age.ind.bc, sd.slice, s, L, L.adj) %>% 
  dplyr::rename(Core = core,
         `Depth [cm]` = depth_cm,
         `$n_{f}$` = n.forams.grp,
         `$n_{rep}$` = n.grps,
         `$\\sigma_{rep}$` = total.age.sd,
         `$SE_{\\sigma_{rep}}$` = sd.of.sd,
         `$\\sigma_{meas}$` = mean.machine.err,
         `$\\sigma_{ind}$` = sd.age.ind,
         `Bias` = bias.correction,
         `$\\sigma_{ind_{adj}}$` = sd.age.ind.bc,
         `$\\sigma_{slice}$` = sd.slice,
         `$L$` = L,
         `$s$` = s,
         `$L_{adj}$` = L.adj) %>% 
  knitr::kable(., "latex", booktabs = TRUE, linesep = "", 
               digits = c(0,1,0,0,0,
                             0,0,0,2,
                             0,0,1,1,1), 
               caption = "Measured standard deviation between replicated 14C measurements on small samples of foraminifera, inferred age-heterogeneity between individual foraminifera and the implied bioturbation depth. Value with the subscript 'adj' have been corrected for bias due to the exponential relationship between radiocarbon activity and age.",
               escape = FALSE) %>% 
  kableExtra::landscape()
```


## Across core figure

```{r sd.age_sed.rate.new.s.2}
## dataframe for contour lines
contour.data <- crossing(
  sed.acc.rate = seq(1, 50, 1),  biot.depth = c(5, 10, 15, 20, 25)) %>%
  mutate(
    sd.age.ind = 1000 * biot.depth / sed.acc.rate,
    bd.lbl = paste0(biot.depth, " cm"),
    bd.lbl = factor(bd.lbl,
      ordered = T,
      levels = c("5 cm", "10 cm", "15 cm", "20 cm", "25 cm")
    )
  )


# Variance estimate from Lougheed paper.
lough <- tibble(sd.age.ind.bc = 4892, s = 2.2, 
                dataset = "Published\ndata",
                age.type = "age.14C.cal",
                site = "T86-10P",
                core = "T86-10P\n(Lougheed et al)",
                n.grps = 1) %>% 
  mutate(L = 1/1000 * s * sd.age.ind.bc,
         L.adj = L)


# Variance estimate from Fagault
# Foraminiferal samples were picked from core 
# MD08-3178 (31°17.09'N/11°29.20'W, 2184 m water depth).
# pers comm Yoann Fagault

fagault <- tibble(sd.age.ind.bc = sqrt(1680^2 - 970^2),
                  s = 12, 
                  dataset = "Published\ndata",
                age.type = "age.14C.cal",
                site = "MD08-3178",
                core = "MD08-3178\n(Fagault et al)",
                n.grps = 1) %>% 
  mutate(L = 1/1000 * s * sd.age.ind.bc,
         L.adj = L)

pub.dat <- bind_rows(lough, fagault)

# Add Lougheed to variance data

C14.sd$dataset <- "This study"

c.dat.2 <- C14.sd %>% 
  filter(#n.grps > 4,
         age.type == "age.14C.cal") %>% 
  bind_rows(., pub.dat) 
  
core.pal <- RColorBrewer::brewer.pal(7, name = "Dark2")
names(core.pal) <- c("GeoB 10054-4", "GeoB 10058-1", "GeoB 10066-7",
                     "OR1-1218-C2-BC", "SO213-84-2", 
                     "T86-10P\n(Lougheed et al)", "MD08-3178\n(Fagault et al)")

p.new.s.2 <- contour.data %>% 
  ggplot(aes(x = sed.acc.rate, y = sd.age.ind, group = biot.depth)) +
  geom_line(linetype = 2) +
  geom_segment(data = c.dat.2, aes(x = s, xend = s,
                                   y = sd.age.ind.bc, yend = sd.age.ind,
                                   group = NULL), colour = "grey40",
               arrow = arrow(length = unit(0.1, "inches"), type = "open"),
               inherit.aes = FALSE) +
  geom_errorbar(data = c.dat.2, aes(x = s,
                                  ymax = sd.age.ind.bc + sd.of.sd.ind,
                                  ymin = sd.age.ind.bc - sd.of.sd.ind, width = 0,
                                colour = core, group = NULL),
                  inherit.aes = FALSE) +
  geom_errorbarh(data = c.dat.2, aes(y = sd.age.ind.bc,
                                   xmax = s + s.err, xmin = s - s.err,
                                   height = 0,
                                   colour = core, group = NULL),
                  inherit.aes = FALSE) +
  geom_point(data = c.dat.2, aes(x = s, y = sd.age.ind.bc,
                                shape = dataset,
                                colour = core, group = NULL),
             size = 2, inherit.aes = FALSE) +
  expand_limits(x = 70) + 
  scale_x_continuous("Sedimentation rate [cm/kyr]", trans = "log10",
                     breaks = c(1, 2, 5, 10, 50)) +
  annotation_logticks(sides = "lb") +
  scale_y_continuous("SD individual age [years]", trans = "log10") +
  scale_shape_manual("", values = c(15, 16)) +
  #scale_color_brewer("", type = "qual", palette = "Dark2", direction = 1) +
  scale_colour_manual("", values = core.pal, breaks = names(core.pal) ) +
  theme_bw() + 
  directlabels::geom_dl(aes(label = bd.lbl), 
                        method = list("last.points", "calc.boxes",
                                      "enlarge.box", box.color = NA,
                                      fill = "white", "draw.rects", cex = 0.8))
```


```{r fig-SD-age-sed-rate, fig.width=7, fig.height=4, fig.cap=""}
p.new.s.2
```

## Implied bioturbation depths

```{r L_summary}
biot.ests <- c.dat.2 %>% 
  ungroup() %>% 
  dplyr::select(dataset, age.type, site, n.grps, s, s.err, L, L.adj) %>% 
  dplyr::filter(complete.cases(s, L)) %>% 
  group_by(dataset, age.type, site, s, s.err) %>% 
  mutate(wt = (1/sqrt(n.grps)) / sum(1/sqrt(n.grps), na.rm = T)) %>% 
  dplyr::summarise(L = sum(L * wt, na.rm = T),
            L.adj = sum(L.adj * wt, na.rm = T)) %>% 
  arrange(desc(dataset, site)) %>% 
  ungroup() %>% 
  select(-dataset)
```


```{r tab-L-summary}
## format for latex
biot.ests %>% 
  ungroup() %>% 
  dplyr::filter(age.type == "age.14C.cal") %>% 
  dplyr::select(-age.type) %>% 
  dplyr::rename(`$SE_s$` = s.err,
         `$L$` = L,
         `$s$` = s,
         `$L_{adj}$` = L.adj) %>% 
  knitr::kable(., "latex", booktabs = TRUE, linesep = "", digits = 1,
               caption = "Implied bioturbation depths",
               escape = FALSE)
```


## Lookup figure for required number of forams for a desired standard deviation of age


```{r lookup_fig}
df.look <- crossing(n.forams = 1:500,
                    sd.gamma = c(1.5, 3, 6, 12, 25, 50, 100, 200, 400, 800, 1600, 3200),
                    #sd.gamma = 2^seq(0, 12, 1),
                    biot.depth = c(5, 10, 15, 20)) %>% 
  mutate(sed.acc.rate = 1000 * biot.depth / (sd.gamma * sqrt(n.forams)),
         lbl = paste0(sd.gamma, " years"))


  nms <- paste0("Assumed bioturbation depth = ",
                unique(df.look$biot.depth), " cm")
  names(nms) <- unique(df.look$biot.depth)
  
  lblr <- as_labeller(nms)
  

p.lookup <- df.look %>% 
  ggplot(aes(x = n.forams, y = sed.acc.rate, colour = lbl, group = sd.gamma)) + 
  geom_line()+
  scale_x_continuous("No. forams in a sample", trans = "log10", limits = c(1, 500),
                     breaks = c(1:3, 5, seq(10, 30, 10), 50, seq(100, 500, 400))) +
  scale_y_continuous("Sediment accumulation rate [cm/kyr]",
                     trans = "log10", limits = c(1, 500),
                     breaks = c(1:4, 5, seq(10, 30, 10), 50, seq(100, 300, 100),  500)) +
  annotation_logticks() + 
  scale_color_grey(start = 0.1, end = 0.1)+
  expand_limits(x = 0) +
  theme_bw() +
  facet_wrap(~biot.depth, ncol = 2,#, label = "label_both", sep = "=")
             label = lblr)

p.lookup <- directlabels::direct.label(p.lookup, list("far.from.others.borders", "calc.boxes",
                                   "enlarge.box", box.color = NA,
                                   fill = "white", "draw.rects"))

```


```{r fig-lookup, fig.height=9, fig.width=7, fig.cap=""}
p.lookup
```

### Subfigure just for L = 10

```{r lookup_fig_10}
p.lookup.10 <- df.look %>% 
  filter(biot.depth == 10) %>% 
  ggplot(aes(x = n.forams, y = sed.acc.rate, colour = lbl, group = sd.gamma)) + 
  geom_line()+
  scale_x_continuous("No. forams in a sample", trans = "log10", limits = c(1, 500),
                     breaks = c(1:3, 5, seq(10, 30, 10), 50, seq(100, 500, 400))) +
  scale_y_continuous("Sediment accumulation rate [cm/kyr]",
                     trans = "log10", limits = c(1, 500),
                     breaks = c(1:3, 5, seq(10, 30, 10), 50, seq(100, 300, 100),  500)) +
  annotation_logticks() + 
  scale_color_grey(start = 0.1, end = 0.1)+
  expand_limits(x = 0) +
  theme_bw() +
  facet_wrap(~biot.depth, ncol = 2,#, label = "label_both", sep = "=")
             label = lblr)

p.lookup.10 <- directlabels::direct.label(p.lookup.10, list("far.from.others.borders", "calc.boxes",
                                   "enlarge.box", box.color = NA,
                                   fill = "white", "draw.rects"))

```


```{r fig-lookup.10, fig.height=4, fig.width=5, fig.cap=""}
p.lookup.10
```



## Downcore SO213-84-2


```{r}
SO213_84_2_downcore <- read.csv("SO213_84_2_downcore.csv")

MUC.14C.dat.2 <- age.var.summary %>% 
  filter(site == "SO213-84-2") %>% 
  dplyr::select(site, sd.age.ind) %>% 
  group_by(site) %>% 
  summarise_if(is.numeric, mean) %>% 
  left_join(SO213_84_2_downcore, .)
```


```{r downcore, fig.height=5.5, fig.width=3}
a <- MUC.14C.dat.2 %>%
  ggplot(aes(x = depth_cm, y = age.14C, group = subcore, colour = subcore)) +
  geom_line() +
  geom_linerange(aes(ymax = age.14C + 2*age.14C.se,
                     ymin = age.14C - 2*age.14C.se),
                 linetype = 1, size = 0.5, show.legend = FALSE, alpha = 0.75) +
  geom_linerange(aes(ymax = age.14C + age.14C.se,
                      ymin = age.14C - age.14C.se),
                  linetype = 1, size = 2, show.legend = FALSE, alpha = 0.75) +
  geom_point(size = 1, colour = "Black") +
  coord_flip() +
  scale_x_reverse("Depth [cm]")+
  theme_bw()+
  theme(panel.grid.minor = element_blank(), legend.position = "top", axis.title.x = element_blank()) +
  expand_limits(y = c(5000, 10000))+
  scale_y_continuous("Radiocarbon age [ka]", breaks = 1000*(6:9), labels = 6:9) +
  scale_color_brewer("", type = "qual", palette = "Dark2")

b <- MUC.14C.dat.2 %>%
  ggplot(aes(x = depth_cm, y = age.14C, group = subcore, colour = subcore)) +
  geom_line() +
  geom_linerange(aes(ymax = age.14C + 2*sqrt(age.14C.se^2 + (sd.age.ind^2)/350),
                      ymin = age.14C - 2*sqrt(age.14C.se^2 + (sd.age.ind^2)/350)),
                 linetype = 1, size = 0.5, show.legend = FALSE, alpha = 0.75) +
  geom_linerange(aes(ymax = age.14C + sqrt(age.14C.se^2 + (sd.age.ind^2)/350),
                      ymin = age.14C - sqrt(age.14C.se^2 + (sd.age.ind^2)/350)),
                 linetype = 1, size = 2, show.legend = FALSE, alpha = 0.75) +
  geom_point(size = 1, colour = "Black") +
  coord_flip() +
  scale_x_reverse("Depth [cm]") +
  expand_limits(y = c(5000, 10000)) +
  theme_bw()  +
  theme(panel.grid.minor = element_blank(), legend.position = "none") +
  scale_y_continuous("Radiocarbon age [ka]", breaks = 1000*(6:9), labels = 6:9)+
  scale_color_brewer("", type = "qual", palette = "Dark2")

p.downcore <- egg::ggarrange(a, b, ncol = 1, labels = c("a", "b"), draw = FALSE)
```


```{r fig-downcore, fig.height=5.5, fig.width=3, fig.cap=""}
p.downcore
```


## SE of SD estimates

```{r fig-SD-n-grps, fig.height=3, fig.width=4, fig.cap="Coefficient of variation for estimates of SD-age or implied bioturbation depth as a function of the number of dated samples."}
dat.sd.n <- crossing(SD = seq(100, length.out = 100),
                n = 2:30) %>% 
  mutate(SE.SD = SDofSD(SD, n),
         SE_SD = SE.SD / SD)

dat.sd.n %>% 
  ggplot(aes(x = n, y = SE_SD)) +
  geom_point(colour = "grey", size = 1) +
  geom_line(alpha = 0.75) +
  expand_limits(y = 0, x = 0) +
  geom_vline(xintercept = c(9), linetype = 3) +
  geom_hline(yintercept = c(1/c(4)), linetype = 3) +
  expand_limits(y = 0, x = 0) +
  labs(y = expression("CV of"~sigma[age]~"or inferred "*italic(L)),
       x = expression("Number of "^"14"*"C measurements")) +
  scale_x_continuous(breaks = c(0, 9, 10, 20, 30),
                     labels = c(0, 9, "", 20, 30)) +
  scale_y_continuous(breaks = c(0, 0.2, 0.25, 0.4, 0.6)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```


## Exponential and Gamma distributions

```{r fig-exp-gamma, fig.width=4.5, fig.height=3, fig.cap="Gamma distribution with shape = 1, 2, 3 and 12."}
exp.gamma <- crossing(age = seq(-1, 40000, length.out = 1000),
                      n.forams = c(1, 2, 3, 12), s = 5, d = 10) %>% 
  mutate(exp.sd = 1e03 * d/s,
         d.gamma = dgamma(age, shape = n.forams, scale = exp.sd/n.forams),
         n.forams.grp = factor(n.forams, ordered = T,
                               levels = rev(unique(n.forams))))

p.exp.gamma <- exp.gamma %>% 
  ggplot(aes(x = age, y = d.gamma, colour = n.forams.grp,
             fill = n.forams.grp)) +
  geom_polygon() +
  geom_vline(aes(xintercept = 1e03 * unique(d) / unique(s)), linetype = 2) +
  scale_colour_viridis_d(expression(italic(n)~"forams"), option = "plasma", direction = -1,
                        guide = guide_legend(reverse = T))+
  scale_fill_viridis_d(expression(italic(n)~"forams"), option = "plasma", direction = -1, 
                       alpha = seq(0.5, 0.2, length.out = 5),
                       guide = guide_legend(reverse = T)) +
# Use coord_cartesian to zoom in on part of the scale without truncating the 
# data used for the polygon
  coord_cartesian(xlim = c(-1, 1e04)) + 
  theme_bw() +
  labs(title = expression(italic(s)~"= 5 cm kyr"^"-1"*"; "*italic(L)*" = 10 cm"),
       x = "Age [years]",
       y = "Probability density") +
  theme(panel.grid.minor = element_blank())
p.exp.gamma
```



